---
title: "Homework #1 for Geospatial Analysis & Remote Sensing (EDS223 - Fall 2026)"
date: 10-10-2026
format:
  pdf:
    toc: false
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
  results: hide
---


```{r}
#| echo: true
#| warning: false
#| message: false
#| results: hide

# Load in Packages 
library(tidyverse)
library(sf)
library(here)
library(tmap)

# Read in geodatabase of EJScreen data at the Census Block Group level
ej <- sf::st_read(here::here(
  "data", "ejscreen", "EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb"))

# Filter dataset to a state
louisiana <- ej |> 
  dplyr::filter(ST_ABBREV == "LA")


# Filter dataset to a county 
east_baton <- louisiana |> 
    dplyr::filter(CNTY_NAME == "East Baton Rouge Parish")


# Find the average values for all variables within counties
ks_counties <- aggregate(louisiana, by = list(louisiana$CNTY_NAME), 
FUN = mean)
```


```{r}
# Make the map output static
tmap_mode("plot")
```


```{r}
#| fig-width: 8
#| fig-height: 7

# Map of people of color

tm_shape(east_baton) + # Specifying dataframe

    tm_polygons(fill = "PEOPCOLORPCT", # People of color variable
      lwd = 0.1, # Reducing line width of block border
      col = "white", # Adding specific color to block border
      fill.scale = tm_scale(values = "brewer.blues"), # Setting color gradient
      fill.legend = tm_legend(title = "People of Color (proportion)", # Adding title to legend
      position = tm_pos_out("right", "center"))) + # Setting legend outside of map

    tm_compass(
      size = 2, text.size = 0.7, position = c("left", "bottom")) + # Adding compass

    tm_scalebar(
      width = 10, text.size = 0.5, position = c("left", "bottom")) + # Adding scalebar

    tm_title( 
    "People of Color\nin East Baton Rouge Parish, LA", # Adding title to map
    size = 1.0, # Customizing size of title text
    position = tm_pos_out("center", "top", pos.h = "left", pos.v = "bottom")) # Setting map title

```



```{r}
#| fig-width: 8
#| fig-height: 7

# Map of the Proximity to RMP (chemical) facilities

tm_shape(east_baton) + # Specifying dataframe

    tm_polygons(fill = "PRMP", # RMP facility proximity variable
      lwd = 0.1, # Reducing line width of block border
      col = "black", # Adding specific color to block border
      fill.scale = tm_scale_intervals(
      style = "quantile",  # Adding quantile breaks
      values = c("yellow", "darkred")), # Setting color gradient
      fill.legend = tm_legend(title = "RMP Facility Proximity\n(facilities/km)", # Adding title to legend
      position = tm_pos_out("right", "center"))) + # Setting legend outside of map

    tm_compass(
      size = 2, text.size = 0.7, position = c("left", "bottom")) + # Adding compass

    tm_scalebar(
      width = 10, text.size = 0.5, position = c("left", "bottom")) + # Adding scalebar

    tm_title(
    "Proximity to Chemical Facilities\nin East Baton Rouge Parish, LA", # Adding title to map
    size = 1.0, # Customizing size of title text
    position = tm_pos_out("center", "top", pos.h = "left", pos.v = "bottom")) # Setting map title


```



## Results

Based on these two maps, we can see there is a pattern between where people of color live and proximity to RMP facilities in East Baton Rouge Parish, Louisiana. According to the EPA, an RMP facility is any stationary industrial source that stores, handles, or processes extremely hazardous toxic substances (EPA, 2026). According to the maps, the block groups with the highest RMP proximity scores are concentrated in the northern part of East Baton Rouge Parish along the Mississippi River. Most of these communities are 80–100% people of color. When we look at the northeastern part of the parish, we can see that the proportion of people of color living there is less than 10%, and these areas are also the farthest from RMP facilities. It is important to note that the RMP variable used in this analysis only measures proximity to the facilities, not actual exposure, and these maps show a correlation rather than a cause. If we were to take this a step further, we could examine the different health problems that these communities deal with. In conclusion, these maps offer a good understanding of how the distribution of people of color in East Baton Rouge Parish varies spatially in relation to proximity to RMP chemical facilities, and they highlight where efforts to mitigate these social inequities could be focused.


### Citation 

Environmental Protection Agency. (2026, August 3). *Risk Management Program (RMP) Rule*. EPA. https://www.epa.gov/rmp 